Gaia DR2 view of the GD-1 stellar stream

### 2018-04-28 paper/gd1.tex

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters are transformed into thin, dynamically-cold streams
of stars that are extremely valuable tracers of the large- and small-scale
properties of mass around the Galaxy.
Most of such streams discovered in the Milky Way reside in its halo, and are therefore primarily sensitive to the distribution of dark matter.
% Many thin stellar streams have been discovered around the Milky Way, the
% majority of which exist in the Galactic halo and are therefore primarily
- cleanest sample, foreground removed w gaia parallaxes; members selected on combination of gaia pm and panstarrs colors
- we thus got the crispest view of the stream to date, revealing in high resolution the rich structure which was only hinted at in the previous generation of data
% We present astrometric data for one of the longest stellar streams --- the
% GD-1 stream --- from the second data release (\DR) of the \gaia\ mission.
% We re-discover significant density variations along the stream by selecting
% candidate member stars using the exquisite \gaia\ data alone, including the
% existence of a prominent, apparent gap in the stream.
% <Using photometric data from the \pans, we refine our selection ...>
These density variations --- and especially the gap --- are possible indications
of past interactions with large, massive perturbers.
<estimate the mass of interaction...>
\end{abstract}

@@

Dynamically cold stellar streams, likely formed from the tidal disruption of globular clusters, are sensitive to the large and small scale distribution of matter in the halo

Thin stellar streams form from the tidal disruption of low-dispersion stellar systems such as globular clusters \citep[e.g.,][]{TODO}. To date, $\approx$13--14 thin streams have been discovered by searching large-area sky surveys such as the Sloan Digital Sky Survey (SDSS) and Pan-STARRS1 survey (PS1). These previously known thin streams span a wide range of lengths \citep[$\sim2$--$60$ deg;][]{bernard14,grillmair06} and masses \citep[$\sim10^3$--$10^5~\msun$;][]{TODO,TODO}, however, only two such streams have clear progenitor systems: NGC 5466 and Palomar 5 \citep{TODO,TODO}. Though it has been shown with many independent methods that modeling these thin streams can place strong constraints on the gravitational potential in which they form \citep{apw14,TODO}, many efforts to apply these models to real streams around the Milky Way have been hampered by the small number of known progenitor systems.

>>>>

\begin{abstract}\noindent % trust me
Tidally-disrupted globular clusters are transformed into thin, dynamically-cold streams
of stars that are extremely valuable tracers of the large- and small-scale
properties of mass around the Galaxy.
Most of such streams discovered in the Milky Way reside in its halo, and are therefore primarily sensitive to the distribution of dark matter.
% Many thin stellar streams have been discovered around the Milky Way, the
% majority of which exist in the Galactic halo and are therefore primarily
- cleanest sample, foreground removed w gaia parallaxes; members selected on combination of gaia pm and panstarrs colors
- we thus got the crispest view of the stream to date, revealing in high resolution the rich structure which was only hinted at in the previous generation of data
% We present astrometric data for one of the longest stellar streams --- the
% GD-1 stream --- from the second data release (\DR) of the \gaia\ mission.
% We re-discover significant density variations along the stream by selecting
% candidate member stars using the exquisite \gaia\ data alone, including the
% existence of a prominent, apparent gap in the stream.
% <Using photometric data from the \pans, we refine our selection ...>
These density variations --- and especially the gap --- are possible indications
of past interactions with large, massive perturbers.
<estimate the mass of interaction...>
% 
% Many thin stellar streams have been discovered around the Milky Way, the
% majority of which exist in the Galactic halo and are therefore primarily
% sensitive to the distribution of dark matter.
% Here we present astrometric data for one of the longest stellar streams --- the
% ``GD-1'' stream --- from data release 2 (\DR) of the \gaia\ mission.
% We re-discover significant density variations along the stream by selecting
% candidate member stars first using only the exquisite \gaia\ data.
% Combined with filtering using Pan-STARRS photometry, we see clear evidence of
% multiple prominent under-densities and haps in the stream.
% These density variations --- and especially the gaps --- are possible
% indications of past interactions with large, massive perturbers.
% Given that the orbit of the stream crosses the Milky Way disk at large radius,
% these perturbers are likely halo objects and would have to have masses XX,
% suggestive of dark matter sub-halos.
\end{abstract}

@@

Dynamically cold stellar streams are formed from the tidal disruption of stellar
systems by the gravitational field of their host galaxy (e.g.,
\citealt{Johnston:1998}).
Over 30 candidate stellar streams have been discovered throughout the halo of
the Milky Way and have a wide range of properties such as length, velocity
dispersion, and density (e.g., \citealt{SoManyPeople,Bonaca:2012}; see also the
recent review \citealt{Grillmair:2006, Newberg:2016}).
Of the thin, likely globular-cluster-origin streams, only two have clear
progenitor systems: NGC 5466 (\citealt{TODO}) and Palomar 5 (\citealt{TODO}).

Stellar streams are

\citealt{Bonaca:2018}

Dynamical modeling of stellar streams or models of mean stream tracks have been used

Though it has been shown with many independent methods that modeling these thin streams can place strong constraints on the gravitational potential in which they form \citep{apw14,TODO}, many efforts to apply these models to real streams around the Milky Way have been hampered by the small number of known progenitor systems.

We revisit...do not try to model or fit for Galactic potential. Focus on stream properties, data.
